---
title: "Meerkat Offspring Updated Final Code"
author: "Caroline L Shearer"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(glmmTMB)
```

IMPORT 1-6 MO OFFSPRING AGGR DATA FROM CSV
```{r}
modeldata2 = read.csv("Offspring_dataset_a.csv", header=T, na.strings=c(""," ","NA","na"))
modeldata2$X3.way<-as.factor(modeldata2$X3.way)
modeldata2$X3.way <- factor(modeldata2$X3.way, levels = c("Dominant control", "Subordinate control", "Dominant treated"))
levels(modeldata2$X3.way)
#0.083 is 5 min (0.083*60=4.98min)
#5661-5062=599
#599/5661=10.6% of observations are less than ~5 min
minlengthmodeldata2 <- subset(modeldata2, obs.time.hr >= 0.083)
```


SUMMARY INFO ABOUT DATA
```{r}
#sex individ count DC
modeldata2.DC.sex<- modeldata2[ which(modeldata2$X3.way%in%c('Dominant control') ), ]
modeldata2.DC.F<- modeldata2.DC.sex[ which(modeldata2.DC.sex$sex%in%c('F') ), ]
length(unique(modeldata2.DC.F$kmp.id.nospace))
modeldata2.DC.M<- modeldata2.DC.sex[ which(modeldata2.DC.sex$sex%in%c('M') ), ]
length(unique(modeldata2.DC.M$kmp.id.nospace))
modeldata2.DC.unk<- modeldata2.DC.sex[ which(modeldata2.DC.sex$sex%in%c('unk') ), ]
length(unique(modeldata2.DC.unk$kmp.id.nospace))
#individ count DC
length(unique(modeldata2.DC.sex$kmp.id.nospace))
#litter count DC
length(unique(modeldata2.DC.sex$litter))

#sex individ count SC
modeldata2.SC.sex<- modeldata2[ which(modeldata2$X3.way%in%c('Subordinate control') ), ]
modeldata2.SC.F<- modeldata2.SC.sex[ which(modeldata2.SC.sex$sex%in%c('F') ), ]
length(unique(modeldata2.SC.F$kmp.id.nospace))
modeldata2.SC.M<- modeldata2.SC.sex[ which(modeldata2.SC.sex$sex%in%c('M') ), ]
length(unique(modeldata2.SC.M$kmp.id.nospace))
modeldata2.SC.unk<- modeldata2.SC.sex[ which(modeldata2.SC.sex$sex%in%c('unk') ), ]
length(unique(modeldata2.SC.unk$kmp.id.nospace))
#individ count SC
length(unique(modeldata2.SC.sex$kmp.id.nospace))
#litter count SC
length(unique(modeldata2.SC.sex$litter))

#sex individ count DT
modeldata2.DT.sex<- modeldata2[ which(modeldata2$X3.way%in%c('Dominant treated') ), ]
modeldata2.DT.F<- modeldata2.DT.sex[ which(modeldata2.DT.sex$sex%in%c('F') ), ]
length(unique(modeldata2.DT.F$kmp.id.nospace))
modeldata2.DT.M<- modeldata2.DT.sex[ which(modeldata2.DT.sex$sex%in%c('M') ), ]
length(unique(modeldata2.DT.M$kmp.id.nospace))
modeldata2.DT.unk<- modeldata2.DT.sex[ which(modeldata2.DT.sex$sex%in%c('unk') ), ]
length(unique(modeldata2.DT.unk$kmp.id.nospace))
#individ count DT
length(unique(modeldata2.DT.sex$kmp.id.nospace))
#litter count DT
length(unique(modeldata2.DT.sex$litter))

#hours observed by maternal treatment
sum(modeldata2.DC.sex$obs.time.hr)
sum(modeldata2.SC.sex$obs.time.hr)
sum(modeldata2.DT.sex$obs.time.hr)
#hours observed total
sum(modeldata2$obs.time.hr)

#total offspring
length(unique(modeldata2$kmp.id.nospace))
#total moms
length(unique(modeldata2$mom.id))
#total litters
length(unique(modeldata2$litter))
#number focals by treatment
table(modeldata2$X3.way)
#number focals by treatment per month (continuous; predetermined binning 0.6-1.5, 1.6-2.5)
table(modeldata2$age.months,modeldata2$X3.way)
#e.g. DC M1: 28+67+101+95+26+16+7
#e.g. DC M2: 3+3+48+158+71+101+25
#total number of focals by month
table(modeldata2$age.months)
#number focals by treatment per location
table(modeldata2$X3.way,modeldata2$location1)
#total number of focals by location
table(modeldata2$location1)

#number focals by treatment per am vs pm
table(modeldata2$X3.way,modeldata2$am.pm)
#total number of focals by am vs pm
table(modeldata2$am.pm)

#mean observation time
mean.time<-mean(modeldata2$obs.time.hr)
60*mean.time

#grp size by maternal treatment group
mean.grp.size<-aggregate(modeldata2$group.size, list(modeldata2$X3.way), mean)
mean.grp.size
#overall mean group size
mean(modeldata2$group.size)
```

AGGRESSION MAM
```{r}
#MAM
week.lit.gs.aggr1to6.off31<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+group.size+Total.monthly.rainfall+sex+location1+am.pm+group.size*Total.monthly.rainfall+X3.way*age.months.cont+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2)

summary(week.lit.gs.aggr1to6.off31)

#MAM according to AIC 
week.lit.gs.aggr1to6.off32<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+ group.size+Total.monthly.rainfall+location1+am.pm+group.size*Total.monthly.rainfall+X3.way*age.months.cont+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off32)

week.lit.gs.aggr1to6.off33<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+ group.size+Total.monthly.rainfall+location1+am.pm+X3.way*age.months.cont+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off33)

week.lit.gs.aggr1to6.off34<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+ group.size+location1+am.pm+X3.way*age.months.cont+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off34)

week.lit.gs.aggr1to6.off35<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+ group.size+location1+X3.way*age.months.cont+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off35)

week.lit.gs.aggr1to6.off36<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+ group.size+location1+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off36)

week.lit.gs.aggr1to6.off37<-glmmTMB(i.aggr.all ~ age.months.cont+ group.size+location1+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off37)

week.lit.gs.aggr1to6.off38<-glmmTMB(i.aggr.all ~ age.months.cont+location1+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off38)

week.lit.gs.aggr1to6.off39<-glmmTMB(i.aggr.all ~ age.months.cont+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)

summary(week.lit.gs.aggr1to6.off39)

week.lit.gs.aggr1to6.off40<-glmmTMB(i.aggr.all ~1+offset(log(obs.time.hr))+(1|mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(week.lit.gs.aggr1to6.off40)

library(bbmle)
AICtab(week.lit.gs.aggr1to6.off31,week.lit.gs.aggr1to6.off32,week.lit.gs.aggr1to6.off33,week.lit.gs.aggr1to6.off34,week.lit.gs.aggr1to6.off35,week.lit.gs.aggr1to6.off36,week.lit.gs.aggr1to6.off37,week.lit.gs.aggr1to6.off38,week.lit.gs.aggr1to6.off39,week.lit.gs.aggr1to6.off40)
```

model diagnostics
```{r}
#look at offset relationship to aggression
timid1<-glmer(i.aggr.all ~ obs.time.hr+(1 |litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
timidplot<-ggpredict(timid1,c("obs.time.hr"),full.data=T)
plot(timidplot,rawdata = T)+scale_y_continuous(limits = c(0, 25))
plot(x=modeldata2$obs.time.hr,y=modeldata2$i.aggr.all)

library(DHARMa)
#week.lit.gs.aggr1to6.off32
simulationOutput <- simulateResiduals(fittedModel = week.lit.gs.aggr1to6.off32, n = 1000)
#shows model doesnt have significant evidence of being zero inflated in nb
testZeroInflation(simulationOutput)
# QQ plot only
plotQQunif(simulationOutput) # left plot in plot.DHARMa()
# residual plot only
plotResiduals(simulationOutput) # right plot in plot.DHARMa()

#plot residuals against particular predictors
#plotResiduals(simulationOutput, YOURPREDICTOR)
plotResiduals(simulationOutput,modeldata2$X3.way)
plotResiduals(simulationOutput,modeldata2$age.months.cont)

#test to see if the simulated dispersion is actually matching the observed dispersion
testDispersion(simulationOutput)

#break it down and look at it by group i.e. what are we breaking it down into in the random effects
#id
simulationOutput = recalculateResiduals(simulationOutput , group = modeldata2$kmp.id.nospace)
testZeroInflation(simulationOutput)
testDispersion(simulationOutput)
# QQ plot only
plotQQunif(simulationOutput) # left plot in plot.DHARMa()
# Residual plot only
plotResiduals(simulationOutput) # right plot in plot.DHARMa()
```

FORWARD STEPWISE PROCEDURE
```{r}

fstep1<-glmmTMB(i.aggr.all ~ 1+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep1)

fstep2<-glmmTMB(i.aggr.all ~ location1 + offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace), data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep2)

fstep3<-glmmTMB(i.aggr.all ~ age.months.cont+location1+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep3)

fstep4<-glmmTMB(i.aggr.all ~ age.months.cont+location1+am.pm+offset(log(obs.time.hr))+(1 |mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep4)

#cant converge with all random eff
fstep5<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+location1+am.pm+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep5)

#cant converge with all random eff
fstep6<-glmmTMB(i.aggr.all ~ X3.way+ age.months.cont+ location1+ am.pm + X3.way*age.months.cont+ offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace),data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep6)

fstep7<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+group.size+location1+ am.pm+X3.way*age.months.cont+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace), data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep7)

fstep8<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+group.size+Total.monthly.rainfall+location1+ am.pm+X3.way*age.months.cont+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace), data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep8)

fstep9<-glmmTMB(i.aggr.all ~ X3.way+age.months.cont+group.size+Total.monthly.rainfall+location1+ am.pm+X3.way*age.months.cont+group.size*Total.monthly.rainfall+offset(log(obs.time.hr))+(1 | mom.id/litter/kmp.id.nospace), data=modeldata2,family=nbinom2, na.action=na.exclude)
summary(fstep9)

AICtab(fstep1,fstep2,fstep3,fstep4,fstep5,fstep6,fstep7,fstep8,fstep9)
```

test for significance likelihood ratio test with x2 distribution
```{r}
library(car)
 Anova(week.lit.gs.aggr1to6.off32) # assesing significance
```

pairwise differences
```{r}
 library(lsmeans)
 ph <- lsmeans(week.lit.gs.aggr1to6.off32, ~ X3.way| age.months.cont)
 summary(as.glht(pairs(ph), by = NULL)) # sig for each pairwise compar.; p-value adjusts, accounting for mult. tests
 plot(ph)
```

PLOTTING
```{r}
library(ggeffects)
library(ggplot2)
library(sjPlot)
library(prediction)
library(lme4)

summary(week.lit.gs.aggr1to6.off32)

#momtrtplot_glm1<-ggpredict(week.lit.gs.aggr1to6.off32,c("age.months.cont","X3.way"),full.data=T)

#predict with time set to 1hr of observation
momtrtplot_glm1_1hr <-ggpredict(week.lit.gs.aggr1to6.off32,c("age.months.cont","X3.way"),condition=c(obs.time.hr=1),full.data=T)
```

Arial font export assistance (since Arial font is not automatically in R)
```{r}
library("ggpubr")
library("grid")
library("extrafont")
fonts()
loadfonts(device = "pdf")
#extrafont::font_import()
```

GLM UPDATED COLOR Fig. 2 
```{r}
#(formatted for digital presentations)
#original colors c("#7570B3", "#E60202", "#FFC107") purple, red, orange 
ggplotfig2color2_no_raw<-ggplot(momtrtplot_glm1_1hr, aes(x, predicted)) + geom_ribbon(aes(ymin=conf.low, ymax=conf.high,fill=group), alpha=0.25)+
    geom_line(aes(color=group),size=1.5) +
     scale_colour_manual(values = c("#3a004d", "#278f98", "#ffca47"))+labs(x=expression('Offspring age (months)'), y="Predicted aggression initiated (per hour)",title = "") +set_theme(axis.textcolor.y = "black",axis.textcolor.x = "black", axis.title.color = "black",base = theme_classic()) + theme(text = element_text(size=20,family="Arial",lineheight=unit(2,"line")), legend.position = c(0.75,0.8),legend.key.width=unit(3.5,"line"),legend.key.size = unit(2,"line"),legend.key = element_rect(fill = "white")) +scale_x_continuous(name="Offspring age (months)", breaks=c(1,2,3,4,5,6)) +scale_fill_manual(values = c("#3a004d", "#278f98", "#ffca47"))
ggplotfig2color2_no_raw<-update_labels(ggplotfig2color2_no_raw, list(fill = "Maternal treatment", colour = "Maternal treatment"))
ggplotfig2color2_no_raw

ggsave(file="figure_2_newcolor_no_raw.pdf",ggplotfig2color2_no_raw,width=9,height=8.5)


#(formatted for publication)
smolggplotfig2color2_no_raw<-ggplot(momtrtplot_glm1_1hr, aes(x, predicted)) + geom_ribbon(aes(ymin=conf.low, ymax=conf.high,fill=group), alpha=0.25)+
    geom_line(aes(color=group),size=0.5) +
     scale_colour_manual(values = c("#3a004d", "#278f98", "#ffca47"))+labs(x=expression('Offspring age (months)'), y="Predicted aggression initiated (per hour)",title = "") +set_theme(axis.textcolor.y = "black",axis.textcolor.x = "black", axis.title.color = "black",base = theme_classic()) + theme(text = element_text(size=10,family="Arial",lineheight=unit(1,"line")), legend.position = c(0.75,0.8),legend.key.width=unit(1.5,"line"),legend.key.size = unit(1,"line"),legend.key = element_rect(fill = "white")) +scale_x_continuous(name="Offspring age (months)", breaks=c(1,2,3,4,5,6)) +scale_fill_manual(values = c("#3a004d", "#278f98", "#ffca47"))
smolggplotfig2color2_no_raw<-update_labels(smolggplotfig2color2_no_raw, list(fill = "Maternal treatment", colour = "Maternal treatment"))
smolggplotfig2color2_no_raw

ggsave(file="smolfigure_2_newcolor_no_raw.pdf",smolggplotfig2color2_no_raw,width=88,height=83, units='mm')
```


GLM B&W Fig. 2
```{r}
#formatted for if B&W needed
ggplotfig2_glm1<-ggplot(momtrtplot_glm1_1hr, aes(x, predicted))+ geom_ribbon(aes(ymin=conf.low, ymax=conf.high,fill=group), alpha=0.55)+
    geom_line(aes(linetype=group, color=group),size=1.5) +
     scale_colour_manual(values = c("black", "black", "black"))+labs(x=expression('Offspring age (months)'), y="Predicted aggression initiated (per hour)",title = "") +set_theme(axis.textcolor.y = "black",axis.textcolor.x = "black", axis.title.color = "black",base = theme_classic()) + theme(text = element_text(size=20,family="Arial",lineheight=unit(2,"line")), legend.position = c(0.75,0.8),legend.key.width=unit(3.5,"line"),legend.key.size = unit(2,"line"),legend.key = element_rect(fill = "white")) +scale_x_continuous(name="Offspring age (months)", breaks=c(1,2,3,4,5,6)) + scale_linetype_manual(values=c("Dominant control"="dashed","Subordinate control"="solid","Dominant treated"="dotted"))+scale_fill_manual(values = c("grey80", "grey45", "grey20"))+ guides(color = guide_legend(override.aes = list(alpha=0.8) ) )
ggplotfig2_glm1<-update_labels(ggplotfig2_glm1, list(shape = "Maternal treatment",fill = "Maternal treatment",linetype = "Maternal treatment",colour = "Maternal treatment"))
ggplotfig2_glm1
ggsave(file="glm_version_no_data_points_figure_2_BW.jpeg",ggplotfig2_glm1,width=9,height=8.5,dpi=600)
```

extraction of DATA for source data file (Supp. Mat. figure s3)
```{r}
#subsetting for source data file
suppmattfig2= minlengthmodeldata2[,c('age.months.cont','i.aggr.all','obs.time.hr','X3.way')]
write.csv(suppmattfig2,'sourcedata_supmattfigure2.csv')
DTsuppmattfig2<- minlengthmodeldata2[ which(minlengthmodeldata2$X3.way%in%c('Dominant treated') ), ]
DTsuppmattfig2= DTsuppmattfig2[,c('age.months.cont','i.aggr.all','obs.time.hr','X3.way')]
DCsuppmattfig2<- minlengthmodeldata2[ which(minlengthmodeldata2$X3.way%in%c('Dominant control') ), ]
DCsuppmattfig2= DCsuppmattfig2[,c('age.months.cont','i.aggr.all','obs.time.hr','X3.way')]
SCsuppmattfig2<- minlengthmodeldata2[ which(minlengthmodeldata2$X3.way%in%c('Subordinate control') ), ]
SCsuppmattfig2= SCsuppmattfig2[,c('age.months.cont','i.aggr.all','obs.time.hr','X3.way')]

#source data file extraction
write.csv(DCsuppmattfig2,'SupplementaryFigure2_SourceData_bytrt.csv')
write.csv(DTsuppmattfig2,'DTSupplementaryFigure2_SourceData.csv')
write.csv(SCsuppmattfig2,'SCSupplementaryFigure2_SourceData.csv')
```


Color Fig. S3 raw data (truncated to focals >5 min [>0.083] to reduce outliers on plot)
```{r}
#formatted for supplementary material
Hexagonal_binning<-ggplot(minlengthmodeldata2, aes(x = age.months.cont, y = i.aggr.all/obs.time.hr,colour=X3.way))+geom_hex(bins=20)+scale_fill_gradientn(colours = c("#f5f4f4","grey","#242526"), name = "Frequency")+scale_colour_manual(values = c("#3a004d", "#278f98", "#ffca47"))+facet_wrap(~X3.way, labeller = labeller(X3.way = c("Dominant control" = "a","Subordinate control" = "b", "Dominant treated" = "c")))+labs(x=expression('Offspring age (months)'), y="Aggression initiated (per hour)",title = "")+theme(plot.margin=unit(c(0,0,0,1.5),"mm"),strip.background = element_blank(),strip.text.x = element_text(hjust = -0.005,size = 15),text = element_text(size=15,family="Arial",lineheight=unit(1.5,"line")),legend.key.width=unit(1.5,"line"),legend.key.size = unit(1.5,"line"),legend.position ="right" )+guides(fill=guide_colorbar(reverse = TRUE),color=guide_legend(order=2,override.aes=list(fill="white",size = 1)),size = "legend")#+theme(strip.background = element_blank(), strip.text = element_blank())
Hexagonal_binning<-update_labels(Hexagonal_binning, list(colour = "Maternal treatment"))
Hexagonal_binning

ggsave(file="m.7_mindata_supp_hex_newcolor_BW.pdf",Hexagonal_binning,width=12,height=4.5,dpi=600)
ggsave(file="m.7_mindata_supp_hex_newcolor_BW.png",Hexagonal_binning,width=12,height=4.5,dpi=300)
```

B&W Fig. S3 raw data (truncated to focals >5 min [>0.083] to reduce outliers on plot)
```{r}
#formatted for if B&W needed
Hexagonal_binning<-ggplot(minlengthmodeldata2, aes(x = age.months.cont, y = i.aggr.all/obs.time.hr,colour="black"))+geom_hex(bins=20)+scale_fill_gradientn(colours = c("#f5f4f4","darkgrey","black"), name = "Frequency")+scale_colour_manual(values = "black")+facet_wrap(~X3.way)+labs(x=expression('Offspring age (months)'), y="Aggression initiated (per hour)",title = "")+ set_theme(axis.textcolor.y = "black",axis.textcolor.x = "black", axis.title.color = "black",base = theme_classic())+theme(plot.margin=unit(c(0,0,0,1.5),"mm"),strip.text.x = element_text(size = 15),text = element_text(size=15,family="Arial",lineheight=unit(1.5,"line")),legend.key.width=unit(1.5,"line"),legend.key.size = unit(1.5,"line"),legend.position ="right" )+guides(fill=guide_colorbar(reverse = TRUE),size = "legend", colour = "none")#+theme(strip.background = element_blank(), strip.text = element_blank())
Hexagonal_binning
ggsave(file="m.6_minlengthdata_supp_hex_BW.jpeg",Hexagonal_binning,width=11,height=4.5,dpi=600)
```
